---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Colin Li"
date: "02/07/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r}

dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

```

How many Denny’s locations are there in Alaska?
3 locations

### Exercise 2

```{r}

lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)

```
How many La Quinta locations are there in Alaska?

2 locations

### Exercise 3

```{r}

dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak


```

...

### Exercise 4


```{r}

glimpse(dn_lq_ak)

```
How many observations are in the joined dn_lq_ak data frame? What are the names of the variables in this data frame.

6 observations, names are address.x, city.x, state, zip.x, longitude.x, latitude.x, country.x, establishment.x, address.y, city.y, zip.y, longitude.y, latitude.y, country.y, establishment.y


...

### Exercise 5

```{r}

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}

```


...

### Exercise 6

```{r}

dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(long1 = longitude.x, lat1 = latitude.x, long2 = longitude.y, lat2 = latitude.y))

```


...

### Exercise 7

```{r}

dn_lq_ak_mindist <- dn_lq_ak %>%
  dplyr::group_by(address.x) %>%
  dplyr::summarize(closest = min(distance))

```

### Exercise 8

```{r}
library(ggsci)

ggplot(data = dn_lq_ak_mindist, 
       mapping = aes(x = closest, color = " ", show.legend = FALSE)) + scale_color_aaas() + geom_freqpoly() + scale_y_continuous(breaks = c(0, 1, 2)) + scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6)) + xlab("Distance (miles)") + ylab("Number of Establishments") + ggtitle("AK: Distances between Denny’s and the Nearest La Quinta") + labs(caption = "There are 3 Denny's-La Quinta pairings within 6 miles") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
) 

summary(dn_lq_ak_mindist)

```

### Exercise 9

```{r nc_graph}

dn_nc <- dennys %>%
  filter(state == "NC")

lq_nc <- laquinta %>%
  filter(state == "NC")

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(long1 = longitude.x, lat1 = latitude.x, long2 = longitude.y, lat2 = latitude.y))

dn_lq_nc_mindist <- dn_lq_nc %>%
  dplyr::group_by(address.x) %>%
  dplyr::summarize(closest = min(distance))


ggplot(data = dn_lq_nc_mindist, 
       mapping = aes(x = closest, color = " ", show.legend = FALSE)) + scale_color_aaas() + geom_freqpoly() + xlab("Distance (miles)") + ylab("Number of Establishments") + ggtitle("NC: Distances between Denny’s and the Nearest La Quinta") + labs(caption = "There are 28 Denny's-La Quinta pairings within 200 miles") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
) 

summary(dn_lq_nc_mindist)

```


### Exercise 10

```{r tx_graph}

dn_tx <- dennys %>%
  filter(state == "TX")

lq_tx <- laquinta %>%
  filter(state == "TX")

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx

dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(long1 = longitude.x, lat1 = latitude.x, long2 = longitude.y, lat2 = latitude.y))

dn_lq_tx_mindist <- dn_lq_tx %>%
  dplyr::group_by(address.x) %>%
  dplyr::summarize(closest = min(distance))


ggplot(data = dn_lq_tx_mindist, 
       mapping = aes(x = closest, color = " ", show.legend = FALSE)) + scale_color_aaas() + geom_freqpoly() + xlab("Distance (miles)") + ylab("Number of Establishments") + ggtitle("TX: Distances between Denny’s and the Nearest La Quinta") + labs(caption = "There are 200 Denny's-La Quinta pairings within 60 miles") +
  theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
) 

summary(dn_lq_tx_mindist)


```

